---
title: "Assignment 1 - Carseats Dataset Analysis"
author: "Kiersten Sweet"
output: html_document
---

# Introduction

This assignment demonstrates importing a dataset, generating descriptive statistics, transforming variables, and creating visualizations using R.

```{r}
library(ISLR)

data(Carseats)

head(Carseats)
```

# Descriptive Statistics

```{r}
summary(Carseats[, c("Sales", "Price", "Income", "Advertising")])

table(Carseats$ShelveLoc)
table(Carseats$Urban)
table(Carseats$US)
```

The summary above shows descriptive statistics for quantitative and categorical variables.

# Variable Transformation

```{r}
Carseats$HighSales <- ifelse(Carseats$Sales > 8, "High", "Low")

table(Carseats$HighSales)
```

A new variable called HighSales was created.

# Histogram

```{r}
hist(Carseats$Sales,
     main = "Distribution of Sales",
     xlab = "Sales")
```

# Scatterplot

```{r}
plot(Carseats$Price,
     Carseats$Sales,
     main = "Price vs Sales",
     xlab = "Price",
     ylab = "Sales")
```

The scatterplot shows the relationship between price and sales.